Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-12-02 inst/paper/main.tex
delete

\subsubsection{Usage Patterns}

Table~\ref{table:js-pattens-in-r} shows the ratio of \eval in each of
these patterns. We included all the original patterns except for
\emph{JSON} as R uses its own serialization format and for
\emph{empty} as there were no empty expression passed to \eval:

\begin{table}[ht]%\resizebox{\columnwidth}{!}{%
  \centering
\begin{tabular}{r|r|r|r|r}\hline
\multirow{2}{*}{Pattern} & \multicolumn{2}{c|}{Core} & \multicolumn{2}{c}{Packages} \\
                       & Calls & Ratio            & Calls & Ratio \\\hline
\input{tag/table-js-patterns-in-r.tex}
\end{tabular}
\caption{JavaScript \eval Patterns in R}
\label{table:js-pattens-in-r}
\end{table}

\noindent \emph{Library}, \emph{Typeof} and \emph{Try} are all barely
used mostly because the language provides dedicated features for it.
The few calls in Library pattern come mostly from file loading
functions such as \c{source} and \c{sys.source}.
%
\noindent
\emph{Read} pattern in JavaScript represented mostly an incorrect
access to an object's property or local variable both replaceable by
JavaScript core functions. In the case of R there are \PatternReadRnd
read \eval calls (\PatternReadRatio). They come from calls like
\c{subset(airquality, select = Temp)} where the variable \c{Temp} will
be evaluated in the \c{airquality} data frame.

%
\noindent \emph{Assign} category comprises of assignments to variables
within an \eval call. We have seen \PatternAssignRnd such calls
(\PatternAssignRatio). \PatternAssignArrowRatio uses the simple
assignment (\c{<-}). The majority of cases comes from the
\c{magrittr}'s \c{\%<>\%} pipe operator
(\PatternAssignArrowMagrittrRatio) which forwards an object to a
function on rhs and update lhs with the result. Far more calls used
the super assignment (\c{<<-}) operator, but almost all come from the
\c{plyr} package which provides a convenient API for common data
manipulation problems.
%
\noindent \emph{Call} and \emph{Other} together represents the vast
majority of eval calls in R. The \emph{Other} pattern represents all
the calls where eval was called with a value instead of with an
expression. The \emph{Call} pattern contains all the calls that did
not fall in the previous categories. Without a context it is usually
difficult to further categorize the intent of these calls. Therefore
we have randomly chosen \PatternManualPackages packages and manually
looked at \PatternManualCallsites call sites from which we have
observed the following categories:

\subsubsection{Implementation}
The package installation process adds an extra file to the package
directory that contains the code to load the package code inside a
\local block. It creates a namespace object and injects the package
bindings.

\subsubsection{Miscellaneous}
\eval is also used in R packages to get around various limitations.
For example, in \datatable package, \eval is used to modify a binding
in the \base package of R. Bindings in the \base package are locked
for modification and require explicit unlocking using function
\unlockBinding. However, directly using this function triggers a
restricted-function-use warning during package build, preventing the
package to be accepted by CRAN. The \datatable gets around this
limitation by enclosing the call to \unlockBinding inside \eval. From
the same reason, other packages use \eval to access private variables
and functions from their dependencies.

\subsubsection{Provenance}

There are various ways to obtain expressions:
\begin{itemize}
  \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
  \item \c{expression}  creates a vector of expression
    objects from text.
  \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
\end{itemize}

@@

\subsubsection{Scope}

One of the main differences of the R \eval with the javascript one is
how it interacts with \emph{scope}, or environments in R. In
Javascript, \eval can access the local scope and the global scope. In
R, \eval has an argument \c{envir} to indicate in which environment
its first argument must be evaluated. By default, it is equal to
\c{parent.frame()}, which is the parent environment of \eval call.
This default argument happens in \DefaultEnvirExprPercent of the eval
calls.

This argument can also be \c{NULL}, a \c{List} or a data frame. This
happens in 3.5\% of the cases: \eval copies the fields of the list or
data frame and creates bindings for them in a new environment. This
pattern is used to evaluate formulas which can directly refer to the
fields of the data. The \c{envir} argument can also be a number $n$.
It means that the environment in which the expressions is evaluated
will be the result of \c{sys.call(n)} where $n$ refers to the $n$-th
stack frame.

@@

\section{Comparison with Javascript}

Comparing the use of \eval in R and JavaScript is instructive. The use
of \eval is more widespread in data science code than it was in
JavaScript. Any run of an R program seems to trigger \eval; moreover
package programmers use \eval liberally. On the other hand, the
authors of Kaggle packages do not use the feature at all. This point
to a bifurcated user community: experts who write package know how to
use \eval while end-users do not. An alternative explanation is that
end-users do not need \eval for simple data analysis tasks. This
bifurcation was not observed in the JavaScript community, but this may
have changed with the advent of Node.js ecosystem. The distribution of
input sizes is similar to that found in JavaScript, the majority of
inputs being rather small.

>>>>

\myoara{Source.}
There are various ways to obtain expressions:
\begin{itemize}
  \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
  \item \c{expression}  creates a vector of expression
    objects from text.
  \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
\end{itemize}

@@

\mypara{Scope.} The environment argument can be \c{NULL}, a \c{List}
or a data frame. This happens in 3.5\% of the cases: \eval copies the
fields of the list or data frame and creates bindings for them in a
new environment. This pattern is used to evaluate formulas which can
directly refer to the fields of the data. The \c{envir} argument can
also be a number $n$. It means that the environment in which the
expressions is evaluated will be the result of \c{sys.call(n)} where
$n$ refers to the $n$-th stack frame.

@@

\subsection{Comparison with Javascript}
